---
title: "Exercise 3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Oussama Mesbah (12100912)"
date: "`r Sys.Date()`"
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).

Done

b)  Commit the files generated by Rstudio.

Done

c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)

Strengths: 
1. Git is a distributed version control system, which means that the entire codebase and history is available, which makes collaboration easy.
2. Git is fast and efficient, as it only stores the changes made to the codebase, not the entire codebase.

Weaknesses:
1. Gits Lack of a GUI makes it hard to use in collaboration with non-technical people.
2. Git has a lot of commands, which makes it hard to learn.

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.
    
Done

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:
    
Github Repository: https://github.com/OussamaMesbah/exeRcise-sheet-3

b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.

Done    

c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.
    
Done

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.

```{r message=FALSE, warning=FALSE, echo=TRUE}
library(readr)
library(dplyr)

names_2021 <- read_csv("vornamen_2021.csv")
names_2022 <- read_csv("vornamen_2022.csv")

typeof(names_2021$Anzahl)
typeof(names_2022$Anzahl)
```
```{r echo=TRUE}
names_2021 %>% slice(1000:1005)
names_2022 %>% slice(1000:1005)
```
As we can see above the issue is that some entries have the qualifier "less than"
for rare names. So the character type is actually appropriate.

We will change these entries with 2, as this is the middle between 1 and 4.This way
some names will be counted too often and others too seldom. This is both a solution 
and the repercussions as we generate some inaccuracy. 

```{r echo=TRUE}
#We change the column "Anzahl" with our new rule
names_2021$Anzahl[names_2021$Anzahl == "4 oder weniger"] <- 2
names_2022$Anzahl[names_2022$Anzahl == "4 oder weniger"] <- 2

# change type to numeric
names_2021$Anzahl <- as.numeric(names_2021$Anzahl)
names_2022$Anzahl <- as.numeric(names_2022$Anzahl)
```

c)  Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
    

```{r echo=TRUE}
sum(names_2021$Anzahl)
sum(names_2022$Anzahl)
```
We can see that the number of newborn was greater in 2021. 

d)  Add a new column `year` to both datasets which holds the correct
    year for each.

```{r}
names_2021$year <- 2021
names_2022$year <- 2022
```

e)  Combine both datasets into one using `bind_rows()`.

```{r}
names_2021_2022 <- bind_rows(names_2021, names_2022)
```

f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.

```{r}
library(knitr)
names_2021_2022 %>% 
  group_by(Vorname) %>% 
  summarise(Anzahl = sum(Anzahl)) %>% 
  arrange(desc(Anzahl)) %>% 
  slice(1:10) %>% 
  kable(caption = "Top 10 names in 2021 and 2022")
```

## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

a)  The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

The following prompt was used:

<blockquote>
The following code does not work. It uses the apply function wrong. change it to the correct apply function and explain why it did not work.  Output the rowwise means.

###Create a sample data frame

tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )

###Calculate column-wise means

column_means \<- lapply(tax_data\\[, -1\], 2, mean)

column_means
</blockquote>

ChatGPT generated the following:

<blockquote>
The code you provided is attempting to calculate the column-wise means using the `lapply` function. However, the usage of `lapply` is incorrect in this context. The `lapply` function is used to apply a function to each element of a list or a vector. In this case, you are trying to calculate the means of each column of the `tax_data` data frame, so `lapply` is not the appropriate function to use.

To calculate the column-wise means, you can use the `colMeans` function instead. Here's the corrected code:

```{r}
# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means
column_means <- colMeans(tax_data[, -1])

column_means
```

This code will correctly calculate the means of each column except for the "Name" column because it is excluded using `tax_data[, -1]`. The resulting `column_means` vector will contain the mean values for "Tax_2019", "Tax_2020", and "Tax_2021".

If you want to calculate the row-wise means instead, you can use the `rowMeans` function. Here's the modified code:

```{r}
# Calculate row-wise means
row_means <- rowMeans(tax_data[, -1])

row_means
```

This will calculate the means for each row of the `tax_data` data frame, excluding the "Name" column. The resulting `row_means` vector will contain the mean values for each row.
</blockquote>



b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
The following prompt was used:

> Explain to me what the rapply() function in R does. For this purpose, create an easy example with some mock data where the function is used and 
explain it.

ChatGPT generated the following:
<blockquote>
The `rapply()` function in R is a recursive version of the `lapply()` function, used to apply a given function to elements of a list or nested lists. It traverses through a list and its nested lists, applying a specified function to each element.

To illustrate the usage of `rapply()`, let's create a simple example. Suppose we have a list called `my_list` that contains some numeric values and nested lists:

```{r}
my_list <- list(1, 2, list(3, 4), list(5, list(6, 7)))
```

In this example, `my_list` contains four elements: two numeric values (1 and 2), a list with two numeric values (3 and 4), and another list with one numeric value (5) and a nested list containing two numeric values (6 and 7).

Now, let's say we want to square each numeric value in the `my_list` structure. We can achieve this using `rapply()`:

```{r}
squared_list <- rapply(my_list, function(x) x^2, how = "replace")
```

In the above code, we provide `my_list` as the input and specify the function `function(x) x^2` to square each element. The argument `how = "replace"` indicates that we want the resulting list to replace the original values with their squared counterparts.

After executing this code, the `squared_list` will contain the same structure as `my_list`, but with each numeric value squared:

```R
[[1]]
[1] 1

[[2]]
[1] 4

[[3]]
[[3]][[1]]
[1] 9

[[3]][[2]]
[1] 16


[[4]]
[[4]][[1]]
[1] 25

[[4]][[2]]
[[4]][[2]][[1]]
[1] 36

[[4]][[2]][[2]]
[1] 49
```

As you can see, the original numeric values in `my_list` have been squared, resulting in a new list structure with the squared values. Note that the nested lists have also been preserved, maintaining the original structure of `my_list`.

In summary, the `rapply()` function allows us to recursively apply a specified function to each element in a list or nested lists, providing a flexible and convenient way to manipulate complex data structures in R.

</blockquote>


## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.
